Structured Versus Decorated Cospans

### 2020-12-31 cospans.tex
improved introduction to Section 5

Decorated cospans were introduced by Fong \cite{Fong,FongThesis} to describe open systems as cospans equipped with extra data.  They were then applied to open electrical circuits \cite{BF}, Markov processes \cite{BFP}, and chemical reaction networks \cite{BP}.  Unfortunately, some of these applications were marred by technical flaws, which were later fixed using structured cospans \cite{BC}. Here we explain how to also fix them using our new decorated cospans, since they provide another another solution to these problems. Below we compare the two approaches in applications to graphs, electrical circuits, Petri nets, reaction networks and dynamical systems. In many cases, \cref{thm:equiv} shows that the structured and decorated cospan approaches are equivalent: \cref{subsec:graphs,subsec:circuits,subsec:petri} illustrate this. However, in \cref{subsec:petrirates} we show that open dynamical systems can only be treated using decorated cospans.

>>>>

Decorated cospans were introduced by Fong \cite{Fong,FongThesis} to describe open systems as cospans equipped with extra data.  They were then applied to open electrical circuits \cite{BF}, Markov processes \cite{BFP}, and chemical reaction networks \cite{BP}.  Unfortunately, some of these applications were marred by technical flaws, which were later fixed using structured cospans \cite{BC}. Here we explain how to also fix them using our new decorated cospans, since they provide another another solution to these problems. Below we compare the two approaches in applications to graphs, electrical circuits, Petri nets, reaction networks and dynamical systems.

In many cases, \cref{thm:equiv} shows that the structured and decorated cospan approaches are equivalent: \cref{subsec:graphs,subsec:circuits,subsec:petri} illustrate this.   However, in\cref{subsec:petrirates} we describe a map from `open Petri nets with rates' to open dynamical systems.  The former are conveniently described using structured cospans, while the latter can only be treated using decorated cospans.   Here \cref{thm:equiv} serves as a bridge that lets us connect the two formalisms.
